# Basic R functions

::: callout-note
## Functions

-   `library()` - Loads packages into R.
-   [`filter()`](basic_functions_extended.qmd){#Filter}- Allows you to retain/filter rows from a data set depending on the condition you want.
-   [`select()`](basic_functions_extended.qmd){#Select} - Allows you to pick/keep specific columns from a data set.
-   [`count()`](basic_functions_extended.qmd){#count} - Counts the number of occurrences of all distinct elements in a column.
-   [`mutate()`](basic_functions_extended.qmd){#Mutate} - Allows you to create and modify columns from a data set.
-   [`arrange()`](basic_functions_extended.qmd){#Arrange} - Allows for the ordering of the data by different variables.
-   [`group_by()`](basic_functions_extended.qmd){#Group_by} - Uses an existing table to form a grouped table by splitting the data sate based on one or more variables.
-   [`summarise()`](basic_functions_extended.qmd){#Summarise} - Allows you to calculate summary statistics and is usually used on grouped data produced by `group_by()`, as it creates one row for each combination of grouping variables.
-   [`tibble()`](basic_functions_extended.qmd){#Tibble} - Allows you to create new data frames.
-   [`drop_na()`](basic_functions_extended.qmd){#Drop_na} - Allows you to to drop all the rows containing missing values.
:::

------------------------------------------------------------------------

## Joining functions

::: callout-note
## Functions

-   `left_join()` - Keeps all the values from the left table and matching rows from the right table.
-   `right_join()` - Keeps all the values from the right table and matching rows from the left table.
-   `inner_join()` - Keeps all values that have matching keys in both tables.
-   `full_join()` - Keeps all values from both tables.
:::

------------------------------------------------------------------------

## Pivot functions

::: callout-note
## Functions

-   `pivot_longer()` - Allows you to lengthen your data set from a wide format to a long format by reducing the number of columns and increasing the number of rows.
-   `pivot_wider()` - Allows you to widen your data set from a long format to a wide format by reducing the number of rows and increasing the number of columns.
:::

------------------------------------------------------------------------

## Useful operators

`|>` or `%\>%`

::: callout
-   This is a pipe operator that allows you to apply multiple functions to your data set in a chain like sequence.
-   You can type this operator by pressing: **shift + control + M**
:::

Here you can see an example of how to use the pipe operator to chain different functions:

```{r}
#| include: false  
library(tidyverse)
library(here)
rodents <- read_tsv(here("data", "rodents_full.txt"))
```

```{r}
#| eval: false
rodents |> 
  filter(species_id=="NL") |> 
  select(species_id, hindfoot_length, sex)
```

`%in%`

::: callout
-   This is a logical operator that allows you to check if an element from one vector is present in another vector.
:::

`!`

::: callout
-   This is a negation operator, which allows you to tell a function to do the opposite.
:::

This is an example of how you can oppose the filter() in order to filter **out** variables:

```{r}
#| eval: false
rodents |> 
  filter(!species_id=="NL")
```

`+`, `-`, `log()`, `*`, `/`

::: callout
-   These are all mathematical operators.
:::

------------------------------------------------------------------------

## Data overview functions

::: callout-note
## Functions

-   `class()` - It provides just the class of the specified object.
-   `str()` - It shows you the type of content in each column.
-   `glipmse()` - While still detailed, it is easier to glance through than str().
:::

------------------------------------------------------------------------

## Graph functions

`ggplot()` is the function that will allow you to plot your datasets.

Now, in order to get a plot you should think about three main components:

::: {style="text-align: center;"}
**Data + Aesthetics + Geometry**</span>
:::

::: callout
-   **Data** is a data frame you want to use (you can write it inside the ggplot() function of you can directly pipe ggplot() from your dataset.
-   **Aesthetics** is used to indicate x and y variables and it can also be used to control the color of the plot.
    -   `aes()`
-   **Geometry** refers totype of plot you want to make:
    -   `geom_boxplot()`
    -   `geom_point()`
    -   `geom_line()`
    -   `geom_histogram()`
    -   `geom_jitter()`
:::

Here you can see two examples of how to do a plot (one with pipe operator and one without):

```{r}
#| include: false  
library(tidyverse)
library(here) 
rodents <- read_tsv (here("data","rodents_full.txt"))
```

```{r}
#| warning: false
ggplot(rodents) + aes(x= species_id, y=hindfoot_length) + geom_boxplot()

rodents |> 
  ggplot() + aes(x= species_id, y=hindfoot_length) + geom_boxplot()
```

### Further Graph functions

::: callout-note
## Functions

-   `facet_wrap()` - Allows for the visualization of individual categories in their own graph.
-   `labs()` - Change axis labels and legend titles.
-   `scale_x_continous() / scale_y_continous()` - default scales for continuous x and y aesthetics, which can be transform these axes.
-   `theme()` - Modify a single plot's theme.
-   `theme_update()` - Modify the active theme, to affect all subsequent plots.
-   `theme_minimal()` - Applies a simplified theme of the plot, with no background shading.
-   `view()` - Allows you to display your data.
:::
