# Web exercises

all functions:

| function | widget | description |
|:-----------------------|:-----------------------|:-----------------------|
| [`fitb()`](https://psyteachr.github.io/webexercises/reference/fitb.html) | text box | fill-in-the-blank question |
| [`mcq()`](https://psyteachr.github.io/webexercises/reference/mcq.html) | pull-down menu | multiple choice question |
| [`torf()`](https://psyteachr.github.io/webexercises/reference/torf.html) | pull-down menu | TRUE or FALSE question |
| [`longmcq()`](https://psyteachr.github.io/webexercises/reference/longmcq.html) | radio buttons | MCQs with long answers |
| [`hide()`](https://psyteachr.github.io/webexercises/reference/hide.html) and [`unhide()`](https://psyteachr.github.io/webexercises/reference/unhide.html) | button | solution revealed when clicked |

### Let's start with normal Multiple Choice Question (mcq
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(webexercises)
library(tidyverse)
library(here)
rodents_full <- read_tsv(here("data", "rodents_full.txt"))
# style_widgets(
#   incorrect = "#983E82",
#   correct = "#59935B",
#   highlight = "#467AAC"
# )
```


```{r results='asis', echo=FALSE}

opts <- c("install.package", 
            "install.packages", 
            answer = "library", 
            "libraries")
q1 <- mcq(opts)

cat("What function loads a package that is already on your computer?", q1) 

```

### Select section

```{r, results='asis', echo=FALSE}

fitb("select", ignore_case = TRUE)

```

What function to pick columns in a dataset?

```{r, results='asis', echo=FALSE}


fitb("select(x,y)", ignore_case = TRUE)

```

How would you write a code that filters for column 'x' and 'y'?

### filter with AND, OR %in%

::: {.webex-check .webex-box}
filter can be used to group rows: `r torf(TRUE)`

```{r, results='asis', echo = FALSE}
optsf <- c(
   "filter(species_id = 'GM')",
   
   answer = "filter(species_id %in% c(DM, NL)",
   
   "filter(species_id == %in% c(DM, NL)")

cat("Which is the correct form?", longmcq(optsf))
```
:::

### NA


`r hide("Click here to see the solution")`

```{r, echo = FALSE}
"it is all of them" 
```

`r unhide()`
```{r, results='asis', echo=FALSE}
opts_p <- c(
   "drop.na",
   "is.na", 
   "na.rm",
   answer = "All of the above")
cat("Which is the correct form?", longmcq(opts_p))
```
What functions treat NA values?

### Vectors

-   What class would be a Vector with numeric and character values

-   *vectors in R can only hold one type of data*

-   dbl and int are numeric vectors

-   chr is a character vector

-   there is no logical vector in this tibble/dataframe



```         
How would I write a vector that shows the number 28?
```

```{r, echo = TRUE, eval = FALSE, webex.hide = "See a hint"}
a <- 28
```
